---
title: "R Lab 4"
author: "Max Tjen"
date: "7/26"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1:

A random sample of 16 Americans in a city yielded the following data on the number of pounds of beef consumed per year: 118 ,115.125,110, 112, 130,117 ,112 ,115 ,120, 113, 118, 119, 122, 123, 126.  


### a) (4 points)
Construct a quantile-quantile plot for the data to check if the annual beef consumtion in the given city can be modelled by a normal distribution.  

```{r}
data = c(118, 115, 125, 110, 112, 130, 117, 112, 115, 120, 113, 118, 119, 122, 123, 126)
qqnorm(data)
qqline(data) 
```
Ans: The data can be modeled by a normal distribution because the data is mostly along the line of best fit  

### b) (4+2=6 points)  
Construct a two sided 95% confidence interval for the mean beef consumption of the people in the given city and interpret it.  

```{r}
t.test(data, mu = 0, alternative = "two.sided", conf.level=0.95)
```

Ans: 95% CI = (115.4234, 121.4516)  
Since there is a 95% chance that the random intervals cover the value of μ, we expect 95% of the intervals to cover the actual value of μ

